---
title: "Assignment 4 - Text Analytics using quanteda"
author: "Tryphosa Htoisanroi"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
execute:
  echo: true
  warning: false
  message: false
---

# Introduction

This assignment applies text analytics methods using the `quanteda` package in R. The purpose is to analyze political text data, identify similarities and differences across documents, and apply scaling techniques such as Wordfish to compare document positions.

Text analytics is increasingly important in political science, social science, and machine learning because it allows researchers to study large collections of documents systematically.

# Load Required Packages

```{r}
library(quanteda)
library(quanteda.textplots)
library(quanteda.textstats)
library(quanteda.textmodels)
library(ggplot2)
```

# US Presidential Inaugural Speeches

The first part of the assignment analyzes U.S. presidential inaugural speeches included in quanteda.

# Load the Corpus

```{r}
data_corpus_inaugural
```

# Corpus Summary

```{r}
summary(data_corpus_inaugural)
```

# Create Tokens

Tokenization converts text into individual words.

```{r}
toks_inaugural <- tokens(
  data_corpus_inaugural,
  remove_punct = TRUE,
  remove_symbols = TRUE,
  remove_numbers = TRUE
)
```

# Remove Stopwords

Stopwords are common words such as “the” and “and” that usually provide little analytical value.

```{r}
toks_inaugural <- tokens_remove(
  toks_inaugural,
  pattern = stopwords("en")
)
```

# Create Document Feature Matrix

```{r}
dfm_inaugural <- dfm(toks_inaugural)

dfm_inaugural
```

# Most Frequent Words

```{r}
freq_words <- textstat_frequency(
  dfm_inaugural,
  n = 20
)

freq_words
```

# Frequency Plot

```{r}
ggplot(
  freq_words,
  aes(
    x = reorder(feature, frequency),
    y = frequency
  )
) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Most Frequent Words in Presidential Speeches",
    x = "Words",
    y = "Frequency"
  )
```

# Similarities and Differences Across Presidents

## Similarities

Several themes consistently appear across presidential inaugural speeches.

Common themes include:

- Democracy
- Freedom
- National unity
- Economic prosperity
- American values

These themes reflect the traditional goals of presidential speeches.

## Differences

Important differences can also be observed across presidents and historical periods.

### Earlier Presidents

Earlier speeches often focused on:

- Nation building
- Constitutional principles
- Territorial expansion

### War-Time Presidents

War-time speeches emphasized:

- Military strength
- National security
- Patriotism

### Modern Presidents

Modern speeches discuss:

- Globalization
- Technology
- Climate change
- Diversity
- International cooperation

Political ideology also affects speech content and language choices.

# Wordcloud Visualization

```{r}
textplot_wordcloud(
  dfm_inaugural,
  max_words = 100
)
```

# What is Wordfish?

Wordfish is a statistical text scaling model used to estimate the positions of documents based on word frequencies.

It is commonly used in:

- Political science
- Policy analysis
- Ideological comparison
- Legislative studies

Wordfish assumes that documents with similar word usage patterns are positioned closer together.

# Run Wordfish Model

```{r}
wf_model <- textmodel_wordfish(dfm_inaugural)

summary(wf_model)
```

# Plot Wordfish Results

```{r}
textplot_scale1d(wf_model)
```

# Interpretation of Wordfish Results

The Wordfish model estimates latent political or rhetorical dimensions across speeches.

Documents positioned closer together share similar language patterns, while documents farther apart differ in emphasis and rhetorical style.

Possible interpretations include:

- Liberal versus conservative language
- Domestic versus foreign policy focus
- War-time versus peace-time rhetoric

# Comparing Positions Using Scaling Methods

Scaling methods help compare positions across documents.

Common methods include:

- Wordfish
- Correspondence analysis
- Topic modeling
- Sentiment analysis

These methods are useful for understanding ideological differences and policy emphasis.

# US Department of Defense Reports

The assignment also requests text analytics on Department of Defense reports.

# Example Workflow

```{r}
# Example structure only

# library(readtext)

# dod_text <- readtext("USDOD/*.txt")

# dod_corpus <- corpus(dod_text)

# dod_tokens <- tokens(
#   dod_corpus,
#   remove_punct = TRUE,
#   remove_numbers = TRUE
# )

# dod_dfm <- dfm(dod_tokens)
```

# Possible Analyses for DOD Reports

Potential analyses include:

- Most frequent defense-related terms
- Security and military emphasis
- Changes in strategic priorities over time
- International relations topics
- Threat analysis language

# APSA eJobs Hackathon

The final section involves creating a corpus from APSA eJobs postings.

# Create Corpus

```{r}
# Example structure

# jobs_text <- readtext("jobs/*.txt")

# jobs_corpus <- corpus(jobs_text)
```

# Potential Text Analytics

Possible analyses include:

- Most requested research skills
- Common teaching fields
- Frequently requested methodologies
- Geographic hiring trends
- Growth of quantitative methods
- AI and machine learning demand

# Example DFM for Job Ads

```{r}
# Example workflow

# jobs_tokens <- tokens(
#   jobs_corpus,
#   remove_punct = TRUE
# )

# jobs_dfm <- dfm(jobs_tokens)
```

# Discussion

This assignment demonstrates how text analytics can be used to study political language, institutional reports, and job market trends.

Quanteda provides powerful tools for:

- Tokenization
- Frequency analysis
- Visualization
- Statistical scaling
- Corpus analysis

The Wordfish model is especially useful for comparing ideological or rhetorical positions across documents.

# Conclusion

This assignment applied text analytics methods using quanteda to analyze presidential inaugural speeches and other political documents.

The analysis identified important similarities and differences across speeches, demonstrated visualization methods, and introduced statistical scaling using Wordfish.

Text analytics provides researchers with powerful tools for studying political communication, institutional language, and social science data.

Overall, this assignment provided valuable experience with corpus analysis, document feature matrices, and political text scaling methods in R.

# References

- Benoit, K. et al. *quanteda: An R package for the quantitative analysis of textual data*
- James, G. et al. *Introduction to Statistical Learning*
- EPPS6323 Knowledge Mining Course Materials